### 2017-01-14 sec-results.tex
Add citation in results section; add explicit probability values.

To visualize these Markov models that represent our latent states, we plot
them as a directed graph where we set the size of a node to be proportional
to its personalized Pagerank score~\cite{Page:1999:PageRank, Jeh:2003:WWW}
where the personalization vector is the initial state distribution for the
Markov model. We let the thickness of a directed edge $(u, v)$ reflect the
probability of taking that edge given that a random walk is currently at
note $u$ (as indicated by the transition matrix). The plots were created
using python-igraph\footnote{\url{http://igraph.org/python/}}.

@@

Figure~\ref{fig:states} includes two such representations we learned. The
first corresponds to a ``quiz taking'' state whereas the second corresponds
to a ``lecture viewing'' state. Clearly, our unsupervised method can
uncover states that do indeed correspond to student behavior modes that we
would expect to find a priori.

We also argue that it is important that the latent state representation be
a Markov model rather than just a discrete distribution over actions in
$\mathbf{A}$ (as would be the case for a traditional single-layer HMM). We
can observe why if we take a closer look at each of the two latent state
representations in Figure~\ref{fig:states} and look at their forum
component (bottom right). We can see that the relative probability of the
forum activities is roughly the same between these states, but the
\emph{transitions} are quite different. In
Figure~\ref{fig:practice-quiz-state} we have a relatively low probability
of walking from the ``forum thread view'' action back to the ``forum thread
list'' action, but in Figure~\ref{fig:lecture-viewing-state} we actually
observe a very strong link in this direction. This difference highlights an
important distinction between these two latent states: in the first you are
more likely to visit the forum \emph{looking for a particular post}, where
in the second you are more likely to visit the forum \emph{to browse
existing posts}. Thus, capturing the action transition matrix within a
latent state is important for capturing detailed insights involving bigrams
of actions.

We can also use our model for cross-course behavior analysis. In
Figure~\ref{fig:cross-course} we show two latent state representations
learned by a 6-state TL-HMM, one from \textretrieval{} and one from
\sustain{}. These two states were chosen as they are the most similar
across the two courses. However, if we look at the transitions we can see
some important differences. First, in the state from \textretrieval{}, we
can see that the probability of returning to the course wiki after viewing
a lecture is considerably lower than that probability in the state from
\sustain{}. We also notice that the self-loop for staying in a lecture
activity in \textretrieval{} is significantly higher probability than it is
in the state from \sustain{}. This gives us some insight into the lecture
viewing behavior of students in these two MOOCs which can possibly be a
reflection of the course's structure. In \textretrieval{}, students are
likely to view lecture videos in succession directly without first visiting
the course wiki, whereas in \sustain{} students are much more likely to
first return the course wiki before watching the next lecture video. This
observation would be lost if we did not consider the transitions between
the actions within the latent states.

@@

\afterpage{%
\begin{landscape}
\begin{figure*}
  \centering
  \begin{tabular}{cccc}
    \textbf{State 0 (Wiki)} & \textbf{State 1 (Lecture)}
    & \textbf{State 2 (Thread View)} & \textbf{State 3 (Lecture/Wiki)}\\\hline
    \includegraphics[width=0.30\textwidth]{figures/text-2state/state0.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/text-2state/state1.png}\\
    \includegraphics[width=0.30\textwidth]{figures/text-3state/state0.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/text-3state/state1.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/text-3state/state2.png}\\
    \includegraphics[width=0.30\textwidth]{figures/text-4state/state0.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/text-4state/state1.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/text-4state/state2.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/text-4state/state3.png}
  \end{tabular}
  \caption{The evolution of states for increasing $K$ for the
  \protect\textretrieval{} MOOC.\ State ``names'' indicate the highest
  probability action(s) within the state.} % I don't know why \protect is needed
  \label{fig:text-state-evolution}
\end{figure*}
\end{landscape}
}

\afterpage{%
\begin{landscape}
\begin{figure*}
  \centering
  \begin{tabular}{cccc}
    \textbf{State 0 (Wiki)}
    & \textbf{State 1 (Lecture/Wiki)}
    & \textbf{State 2 (Thread List)}
    & \textbf{State 3 (Thread View)}\\\hline
    \includegraphics[width=0.30\textwidth]{figures/sustain-2state/state0.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/sustain-2state/state1.png}\\
    \includegraphics[width=0.30\textwidth]{figures/sustain-3state/state0.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/sustain-3state/state2.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/sustain-3state/state1.png}\\
    \includegraphics[width=0.30\textwidth]{figures/sustain-4state/state0.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/sustain-4state/state3.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/sustain-4state/state1.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/sustain-4state/state2.png}
  \end{tabular}
  \caption{The evolution of states for increasing $K$ for the
  \protect\sustain{} MOOC.\ State ``names'' indicate the highest
  probability action(s) within a state.} % I don't know why \protect is needed
  \label{fig:sustain-state-evolution}
\end{figure*}
\end{landscape}
}

In Figure~\ref{fig:text-state-evolution} we see\footnote{Due to space
constraints, the graphs are quite small, but the states are in the same
positions as they were in previous figures to aid interpretation.} the
evolution of the latent state representations found for the
\textretrieval{} MOOC.  With $K=2$ we uncover a video watching pattern
(state 1) and a course material browsing pattern (state 0). However, we can
see that when $K=3$ we are able to uncover a new pattern involving forum
behavior in state 3 (notice the node sizes on the bottom right). As we
increase $K$ to four, we can see that state 1 splits into state 1 and state
3. These states appear quite similar at a glance, but there are still some
key differences. First, we can see that state 1 now has a non-negligible
forum component, whereas state 3 has hardly any weight on the forum.

@@

In Figure~\ref{fig:trans-low} we plot the latent state transition diagram
for a second group of ``low'' students. These students were selected so
that they attempted all required quizzes in the course, but such that their
average quiz score was $\leq 70\%$. Here, we see that \emph{state 1} has a
large increase in size, where we might have expected state 3 to grow
instead.  However, there is an alternative explanation for this phenomenon:
Since state 1 seems to indicate a highly engaged student, it is a perfectly
reasonable explanation for the ``low'' student group as they are going to
be working hard to try to fill in the gaps in their knowledge. By contrast,
the ``perfect'' student group likely has many members who can take the quiz
more passively and get perfect marks, perhaps because they already know
much of the material being presented, or are just naturally strong and do
not require much background review to perform well. This also explains why
state 1 did not increase in size for the ``perfect'' group like we were
anticipating. Our model lends itself well to discovering this potentially
counter-intuitive insight directly from data.

>>>>

To visualize these Markov models that represent our latent states, we plot
them as a directed graph where we set the size of a node to be proportional
to its personalized Pagerank score~\cite{Page:1999:PageRank, Jeh:2003:WWW}
where the personalization vector is the initial state distribution for the
Markov model. We let the thickness of a directed edge $(u, v)$ reflect the
probability of taking that edge given that a random walk is currently at
note $u$ (as indicated by the transition matrix)\footnote{We do not plot
the transition probabilities directly within the figure to ease
readability; we instead will mention relevant transition probabilities in
the text as we discuss the plots. The plots were created using
python-igraph: \url{http://igraph.org/python/}.}.

@@

Figure~\ref{fig:states} includes two such representations we learned. Under
our interpretation, the first corresponds to a ``quiz taking'' state (it
has higher ``quiz: start'' and ``quiz: submit'' state probabilities than
the other five states) whereas the second corresponds to a ``lecture
viewing'' state. Our unsupervised method can uncover states that do indeed
correspond to student behavior modes that we would expect to find a priori.

We also argue that it is important that the latent state representation be
a Markov model rather than just a discrete distribution over actions in
$\mathbf{A}$ (as would be the case for a traditional single-layer HMM). We
can observe why if we take a closer look at each of the two latent state
representations in Figure~\ref{fig:states} and look at their forum
component (bottom right). We can see that the relative probability of the
forum activities is roughly the same between these states, but the
\emph{transitions} are quite different. In
Figure~\ref{fig:practice-quiz-state} we have a relatively low probability
of walking from the ``forum thread view'' action back to the ``forum thread
list'' action (see the bottom rightmost arc; transition probability $p
\approx 0.17$), but in Figure~\ref{fig:lecture-viewing-state} we actually
observe a much stronger link in this direction (transition probability $p
\approx 0.63$). This difference highlights an important distinction between
these two latent states: in the first you are more likely to visit the
forum \emph{looking for a particular post}, where in the second you are
more likely to visit the forum \emph{to browse existing posts}. Thus,
capturing the action transition matrix within a latent state is important
for capturing detailed insights involving bigrams of actions.

We can also use our model for cross-course behavior analysis. In
Figure~\ref{fig:cross-course} we show two latent state representations
learned by a 6-state TL-HMM, one from \textretrieval{} and one from
\sustain{}. These two states were chosen as they are the most similar
across the two courses. However, if we look at the transitions we can see
some important differences. First, in the state from \textretrieval{}, we
can see that the probability of returning to the course wiki after viewing
a lecture (transition probability $p \approx 0.24$) is considerably lower
than that probability in the state from \sustain{} (transition probability
$p \approx 0.57$). We also notice that the self-loop for staying in a
lecture activity in \textretrieval{} ( transition probability $p \approx
0.70$) is significantly higher probability than it is in the state from
\sustain{} (transition probability $p \approx 0.34$).  This gives us some
insight into the lecture viewing behavior of students in these two MOOCs
which can possibly be a reflection of the course's structure (which, as
demonstrated in \citet{Davis:2016:EDM}, can differ substantially across
different MOOCs).  In \textretrieval{}, students are likely to view lecture
videos in succession directly without first visiting the course wiki,
whereas in \sustain{} students are much more likely to first return the
course wiki before watching the next lecture video. This observation would
be lost if we did not consider the transitions between the actions within
the latent states.

@@

\afterpage{%
\begin{landscape}
\begin{figure*}
  \centering
  \begin{tabular}{cccc}
    \textbf{State 0 (Wiki)} & \textbf{State 1 (Lecture)}
    & \textbf{State 2 (Thread View)} & \textbf{State 3 (Lecture/Wiki)}\\\hline
    \includegraphics[width=0.30\textwidth]{figures/text-2state/state0.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/text-2state/state1.png}\\
    \includegraphics[width=0.30\textwidth]{figures/text-3state/state0.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/text-3state/state1.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/text-3state/state2.png}\\
    \includegraphics[width=0.30\textwidth]{figures/text-4state/state0.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/text-4state/state1.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/text-4state/state2.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/text-4state/state3.png}
  \end{tabular}
  \caption{The evolution of states for increasing $K$ for the
  \protect\textretrieval{} MOOC.\ Each row corresponds to the next value of
  $K$, starting from $K=2$. State ``names'' indicate the highest
  probability action(s) within the state.} % I don't know why \protect is needed
  \label{fig:text-state-evolution}
\end{figure*}
\end{landscape}
}

\afterpage{%
\begin{landscape}
\begin{figure*}
  \centering
  \begin{tabular}{cccc}
    \textbf{State 0 (Wiki)}
    & \textbf{State 1 (Lecture/Wiki)}
    & \textbf{State 2 (Thread List)}
    & \textbf{State 3 (Thread View)}\\\hline
    \includegraphics[width=0.30\textwidth]{figures/sustain-2state/state0.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/sustain-2state/state1.png}\\
    \includegraphics[width=0.30\textwidth]{figures/sustain-3state/state0.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/sustain-3state/state2.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/sustain-3state/state1.png}\\
    \includegraphics[width=0.30\textwidth]{figures/sustain-4state/state0.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/sustain-4state/state3.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/sustain-4state/state1.png}
    &
    \includegraphics[width=0.30\textwidth]{figures/sustain-4state/state2.png}
  \end{tabular}
  \caption{The evolution of states for increasing $K$ for the
  \protect\sustain{} MOOC.\ Each row corresponds to the next value of $K$,
  starting from $K=2$. State ``names'' indicate the highest
  probability action(s) within a state.} % I don't know why \protect is needed
  \label{fig:sustain-state-evolution}
\end{figure*}
\end{landscape}
}

In Figure~\ref{fig:text-state-evolution} we see\footnote{The graphs are
quite small, but the states are in the same positions as they were in
previous figures.} the evolution of the latent state
representations found for the \textretrieval{} MOOC.  With $K=2$ we uncover
a video watching pattern (state 1) and a course material browsing pattern
(state 0). However, we can see that when $K=3$ we are able to uncover a new
pattern involving forum behavior in state 3 (notice the node sizes on the
bottom right). As we increase $K$ to four, we can see that state 1 splits
into state 1 and state 3. These states appear quite similar at a glance,
but there are still some key differences. First, we can see that state 1
now has a non-negligible forum component, whereas state 3 has hardly any
weight on the forum.

@@

In Figure~\ref{fig:trans-low} we plot the latent state transition diagram
for a second group of ``low'' students. These students were selected so
that they attempted all required quizzes in the course, but such that their
average quiz score was $\leq 70\%$. Here, we see that \emph{state 1} has a
large increase in size, where we might have expected state 3 to grow
instead.  However, there is an alternative explanation for this phenomenon.
Since state 1 seems to indicate a highly engaged student, it is a perfectly
reasonable explanation for the ``low'' student group as they are going to
be working hard to try to fill in the gaps in their knowledge. By contrast,
the ``perfect'' student group likely has many members who can take the quiz
more passively and get perfect marks, perhaps because they already know
much of the material being presented, or are just naturally strong and do
not require much background review to perform well. This also explains why
state 1 did not increase in size for the ``perfect'' group like we were
anticipating. \citet{Kizilcec:2017:CandE} observe similar phenomena with
the courses they studied where they find that certificate earning is
negatively correlated with help seeking behavior. Our model lends itself
well to discovering this potentially counter-intuitive insight directly
from data.
